---
title: "5_hub_genes"
author: "Flavio Gabrieli"
format: html
editor: visual
---

# Identification of hub genes, imputation of gene significance, connectivity plot

Resume session

```{r}
session_name <- "PROVA"
```

Load environment

```{r}
source("src/5_env_setup.R")
load(paste0(session_name, "/out/3_networks.Rdata"))
load(paste0(session_name, "/out/1_prepared_data.Rdata"))
source(paste0(session_name, "/in/", session_name,"_parameters.R"))
```

## Hub genes

Load TOM

```{r}
load(net$TOMFiles)
```

2.  Calculate Intramodular Connectivity

    I**ntramodular connectivity** for each gene is calculated as the sum of TOM values for that gene with others in the same module.

```{r}
net_km$IMconnectivity <- intramodularConnectivity(adjMat = TOM, colors = net_km$colors)
rownames(net_km$IMconnectivity) <- colnames(data$exprData)
write.table(net_km$IMconnectivity, file = paste0(session_name,"/out/intramodular_connectivity_km.csv"))

net$IMconnectivity <- intramodularConnectivity(adjMat = TOM, colors = net$colors)
rownames(net$IMconnectivity) <- colnames(data$exprData)
write.table(net$IMconnectivity, file = paste0(session_name,"/out/intramodular_connectivity_hc.csv"))
```

3.  Calculate Module Membership

    Module membership, also called eigengene-based connectivity, measures how strongly a gene correlates with the modules eigengene. Genes with a high absolute value of module membership (e.g., \|MM\| \> 0.8) are considered strongly connected to the module. You can also use signedKME() function

```{r}
net_km$geneMM <- cor(data$exprData, net_km$MEs)
net$geneMM <- cor(data$exprData, net$MEs)
write.table(net_km$geneMM, file = paste0(session_name,"/out/geneMM_km.csv"))
write.table(net$IMconnectivity, file = paste0(session_name,"/out/geneMM_hc.csv"))
```

4.  **Combine Connectivity and Module Membership to identify hub genes**

    Define hub genes as those with:

    -   High intramodular connectivity (e.g., top 10% of kWithin in the module considered).

    -   Strong module membership (e.g., \|MM\| \> 0.8).

```{r}
moduleColor = "skyblue" # select module to analyze

network = "net_km"
assign(paste0(moduleColor, "_",network,"_hubGenes"), hub_genes(network = get(network),
                                                               module = moduleColor,
                                                               minMM = 0.8, # |MM| > 0.8
                                                               q_kWithin = 0.9) # top 10% of kWithin in the module
       )

#export table
write.table(rownames(get(paste0(moduleColor, "_",network,"_hubGenes"))), file = paste0(session_name,"/out/",moduleColor, "_",network,"_hubGenes.txt"), quote = FALSE, col.names = FALSE, row.names = FALSE)
```

## Calculate Gene Significance (GS)

"we define a node significance measure GS = (GS1,\...,GSn) as a vector with n components that correspond to the network nodes. For the ith node, GSi quantifies the significance or importance with regard to a particular application. The only assumption is that GSi = 0 means that node i is not significant with regard to the application under consideration." (Steve Horvath - Weighted Network Analysis\_ Applications in Genomics and Systems Biology (2011, Springer) \[10.1007_978-1-4419-8819-5\])

Using the Gene significance (GS) and module membership (MM) measures, we can identify genes that have a high significance for the trait of interest as well as high module membership in interesting modules.

```{r}
head(data$metadata)

GS <- apply(t(data$exprData), 1, function(gene_expr) cor.test(gene_expr, data$traits$viability, alternative = "two.sided" , method = "pearson")$estimate)
GS_pvalue <- apply(t(data$exprData), 1, function(gene_expr) cor.test(gene_expr, data$traits$viability, alternative = "two.sided" , method = "pearson")$p.value)

max(GS)
```

## Connectivity plot

```{r}
modules = c("skyblue","lightgreen", "cyan", "salmon")

probes = colnames(data$exprData)#get gene names
inModule = is.finite(match(net_km$colors, modules))#get a mask for selecting genes in the
modProbes = probes[inModule]#get the genes
modTOM = TOM[inModule, inModule]#filter the TOM for the genes in the modules of interest
dimnames(modTOM) = list(modProbes, modProbes) #set the names of the modTOM as gene names

# Set the diagonal of the TOM dissimilarity to NA 
dissTOM <- 1-modTOM
diag(dissTOM) <- NA

#hierarchical clustering
geneTree = hclust(as.dist(dissTOM), method="average")

pdf(file = paste0(session_name,"/plots/connectivity_",paste0(modules, collapse = "-"),".pdf"),
  width = 8,          # Width in inches
  height = 8,         # Height in inches
  bg = "white")
TOMplot(dissim = dissTOM^8,
        dendro = geneTree,
        Colors = net_km$colors[inModule],
        main = "Connectivity plot")
dev.off()
```

```{r}
png(file = paste0(session_name,"/plots/connectivity_",paste0(modules, collapse = "-"),".png"),
  width = 8,          # Width in inches
  height = 8,  # Height in inches
  units = "in",
  bg = "white",
  res = 600)
TOMplot(dissim = dissTOM^8,
        dendro = geneTree,
        Colors = net_km$colors[inModule],
        main = "Connectivity plot")
dev.off()
```

## Export network to Cytoscape

```{r}
#indicate one or more modules to export
modules = c("skyblue")

probes = colnames(data$exprData)#get gene names
inModule = is.finite(match(net_km$colors, modules))#get a mask for selecting genes in the indicated modules
modProbes = probes[inModule]#get the genes
modTOM = TOM[inModule, inModule]#filter the TOM for the genes in the modules of interest
dimnames(modTOM) = list(modProbes, modProbes) #set the names of the modTOM as gene names
altNamses = get_TAIRsymbols(modProbes) # get TAIR symbols for annotation of the network
edge_threshold =  calculateTop10PercentThreshold(tom = modTOM)#define a threshold for edges

exportNetworkToCytoscape(modTOM,
  edgeFile = paste("out/cytoscape/CytoscapeInput-edges-", paste(modules, collapse="-"),"edge_threshold", edge_threshold,".txt", sep=""),
  nodeFile = paste("out/cytoscape/CytoscapeInput-nodes-", paste(modules, collapse="-"), "edge_threshold", edge_threshold,".txt", sep=""),
  weighted = TRUE,
  threshold = edge_threshold,
  nodeNames = modProbes,
  altNodeNames = altNamses,
  nodeAttr = bwnet_km$colors[inModule])
```

```{r}
A <-adjacency(data$exprData)

TOMgraph <- graph_from_adjacency_matrix(A[inModule, inModule], mode = "undirected", weighted =TRUE, diag = F)

plot(TOMgraph, vertex.size = 30, vertex.label.color = "white", edge.color = "grey") 
plot(TOMgraph, layout = layout_with_fr(TOMgraph))
```
